Growth & Inequality in Dynamic Public Goods Games

### 2022-10-08 analysis/quarto/paper.qmd
added "quarto" and "paper" (already deprecated) + no clue what happened otherwise. Its been a long time

---
title: "Growth and inequality in public good provision: an extended replication"
description: |
  Report and analysis protocoll.
author:
  - name: Hauke Roggenkamp 
    url: https://www.ibt.unisg.ch/team/hauke-roggenkamp/
    affiliation: Helmut Schmidt University
    affiliation_url: https://www.hsu-hh.de/en/
  - affiliation: Institute for Behavioral Science and Technology
    affiliation_url: https://www.ibt.unisg.ch/
date: "`r Sys.Date()`"
bibliography: ../biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: false
---

@@

@@

> "There are two possible articles you can write: (a) the article you planned to write when you designed your study or (b) the article that makes the most sense now that you have seen the results. They are rarely the same, and the correct answer is (b)." [@bem2021writing, p. 171]

# Introduction

Macroeconomists understand the world as dynamic where the timing matters: Today's savings decision affect lifetime consumption which is why it makes a difference whether you consume more today or tomorrow. Yet, experimental economists often design static games without dynamic inter-dependencies: participants face the same decisions multiple times and previous decisions do not affect current incentives of endowments.

To measure how cooperation affects growth and inequality, @GMTV2017 (to whom I often refer as GMTV in what follows) extended a standard public goods game with temporal inter-dependencies.^[As well as the option to punish group members.] Their results show that participants behave strikingly similar as in the static games (contributions are non-zero; the share of current endowments contributed declines over time; and the initial contribution is around half of their endowments) and illustrate how this affects growth and inequality.

Because environmental and climate action (or the absence of it) is characterized by pro-social behavior and inter-temporal dynamics (the more we hesitate to cut CO2 emissions, the more expensive it will be to reach a certain target in the future), we are interested in applying this setting in a behavioral and environmental economics context with a general population sample.

Due to COVID-19 and corresponding restrictions at that time, we decided to [replicate](#gmtv-replication) GMTV's no punishment 10 period game to evaluate, whether it is feasible to run a multi-period, synchronous group experiment with such a sample first.

As we were not able to recruit the expected number of subjects in the general population sample, we added student participants. Because of this and because the experiment also contains voluntary climate action (VCA), we have a setup similar to @GKLS2020, which allows us to assess [generalizability](#generalizability).

Hence, our paper makes three contributions: First, we run a pre-registered^[The replication project is registered in the AEA RCT Registry and the unique identifying number is: [AEARCTR-0007902](https://doi.org/10.1257/rct.7902-2.0) [@preregistration].] replication of one of GMTV's treatment arms and arrive at similar conclusions: Absolute contributions increase over time. Just as in _static_ public goods experiments, the share of endowments contributed decreases over time. The richest groups earn fifteen times more than the poorest groups. While there clearly is growth, groups do not realize the maximal potential efficiency and earn just over `r main[treatment == "replication" & round == 10, mean(stock)/(80*1.5^10)*100] %>% round(digits=1)`% of what is possible.

Second, like @AGM2018, we study a general population sample online, synchronously and over the course of multiple periods and we find that classic [lab procedures](online-methodology) can be applied outside of the lab too. Relying on an inexperienced sample that cannot interact with experimenters, `r trunc((R1[comprehension > 1] %>% NROW() / R1[!(is.na(comprehension))] %>% NROW())*100)`% of all participants stated that they understood the game. Participants make relatively fast decisions which makes longer games feasible in the future.^[Taken together with inefficient growth and the lack of dropouts, additional rounds are fairly cheap.] In addition, dropouts (or "attrition") are no concern.

Third, we extend the literature on generalizability and find that:

The paper is organized along these lines: Chapter x ...

# Experimental Design and Procedures

This experiment replicated Gächter et al.'s NOPUNISH 10-round treatment arm as close as possible (given the remote circumstances). A demonstration of the experiment can be found [here](https://cliccs.herokuapp.com/).^[Click [here](https://github.com/Howquez/coopUncertainty) to visit the corresponding Github repository.]

All participants were recruited in by the University of Hamburg's  WISO Research Lab using HROOT [@hroot].

@@

## First Round

@@

First, we ask whether the samples differ with respect to their initial contributions to the public good. Is our replication sample more pro-social than the original sample? As Figure 1 reveals, they are not. The distributions look fairly similar and both samples contributed 10 tokens, that is, 50% of their endowments on average (median and mean).^[The two-sided rank sum test (comparing differences between data sources) yields a p-Value of `r rs1` for the mean contribution in first round of the game.] As such, the individuals' behavior in both samples is not only comparable to each other but also to initial contributions in the standard game with partner matching.^[See Figure 3B in @fehrgaechter2000, for instance.] Because participants played a dynamic game, we are particularly interested in the subsequent periods. Do the two groups remain similar?

```{r firstRoundViz}
ggplot(data = R1[treatment != "noPunish15"],
       mapping = aes(x = ownContribution, fill = treatment)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(0, 20),
                       expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, 0.1),
                       expand = c(0, NA)) +
  scale_fill_manual(values = c(cPrimary, cSecondary),
                    labels = c("Original Sample", "Replication Sample")) +
  labs(title = "Individual contributions to the dynamic public good in the first period", 
       y = "", x = "Contributrions (Tokens)") +
  layout +
  theme(legend.position = "bottom")
```

```{r firstRoundSummary, results = "asis", warning = FALSE, fig.cap = "test", eval = FALSE}

@@

We proceed by further discussing contributions over time on a group level. The left panel in Figure 1 shows the average amount of tokens groups contributed over time. Contributions are clearly non-zero and are increasing over the course of the first eight periods. While contributions flatten in the replication, the GMTV data exhibit a notable drop in the last round.

@@

upperLimit <- SUM$contribution %>% max() %>% round() + 10

p1 <- ggplot(data = SUM, 
             aes(x = round, y = contribution, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Average Amount of Tokens contributed") +
  scale_color_manual(values = c(cPrimary, cSecondary)) + 
  theme(plot.margin = margin(0.25,1,0.25,0.25, "cm"))

@@

```{r plotShareOfContributions, fig.cap = "The average amount of tokens contributed over time in treatments."}

@@

p2 <- ggplot(data = SHARE, 
             aes(x = round, y = share, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Share of Current Endowment contributed") +
  scale_color_manual(values = c(cPrimary, cSecondary))

@@

rm(list = c("p1", "p2"))
```

Note that increasing contributions over time imply that groups have increasing endowments over time. Hence, increasing contributions do not necessarily imply that groups contribute increasing shares of their endowments. The right panel in Figure 1 shows the share of overall endowments contributed over time. In the original data, groups contribute around `r SHARE[round == 1 & treatment == "noPunish10", round(share*100)]`% of their endowment in round 1. This amount steadily decreases to `r SHARE[round == 10 & treatment == "noPunish10", round(share*100)]`%. The replication data exhibit a similar pattern with shares ranging from `r SHARE[round == 1 & treatment == "replication", round(share*100)]`% to `r SHARE[round == 10 & treatment == "replication", round(share*100)]`%. `r rm("SHARE")`

Taken together, our replication partly mirrors the results of GMTV. Absolute contributions tended to be higher in @GMTV2017 but end up at around the same level as in the replication due to a stark decline in contributions in the last round. In terms of shares contributed both data sources exhibit a similar pattern: they decline and do not stabilize. Even though the share of current endowments contributed in the last round is quite similar, the share declined a little faster in our data, however.

@@

Possibly of more interest are the implications contributions have for wealth generation and growth. To measure growth, we define a variable called _stock_ which sums the endowments of all participants in a given group at the end of the round (that is, after the contributions have been made, multiplied and redistributed). @GMTV2017 refer to that variable as "wealth" so we will do the same in what follows. Before the start of round 1, wealth will be 80 in all groups by construction. The maximal wealth that can be reached in round 10 (if each group member contributes their entire endowment in each round) is approximately `r (80*1.5^10) %>% trunc()` tokens or `r (80*1.5^10 /20) %>% trunc()` Euro per group.

Table 3 shows some summary statistics regarding wealth at the end of the last period.  While there is clearly growth, groups do not realize the maximal potential efficiency as the replication groups reach on average a level of `r main[treatment == "replication" & round == 10, mean(stock)] %>% trunc()` tokens out of `r (80*1.5^10) %>% trunc()` maximally possible or `r main[treatment == "replication" & round == 10, mean(stock)/(80*1.5^10)*100] %>% round(digits=1)`%. As in the original data, there is large heterogeneity with the richest group reaching `r main[treatment == "replication" & round == 10, max(stock)]` tokens whereas the poorest group ends up with `r main[treatment == "replication" & round == 10, min(stock)]` tokens.

```{r summary1, results = "asis", warning = FALSE, fig.cap = "test"}

@@

stargazer(temp,
          summary.stat = c("mean", "median","sd", "max", "min", "n"),
          type = type, 
          flip = TRUE, 
          header=FALSE)

@@

@@

Figure 2 shows the dynamics of wealth over time. The left panel focuses on all groups, the upper right panel on those with above median wealth after round 10 (“successful” groups) and the lower left panel on those with below median wealth after round 10 (“unsuccessful” groups).

```{r plotStock, fig.cap = "Average wealth over time across treatments."}
STOCK <- main[,
              lapply(.SD, mean, na.rm = TRUE),
              by = c("round", "treatment"),
              .SDcols = "stock"]

@@

p1 <- ggplot(data = STOCK, 
       aes(x = round, y = stock, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          # annotate("segment", x = 1, xend = 10, 
          #          y = STOCK[round == 1 & treatment == "replication", stock],
          #          yend = STOCK[round == 10 & treatment == "replication", stock],
          #          colour = "black", lty = 2, alpha = 0.5) +
          # annotate("segment", x = 1, xend = 10, 
          #          y = STOCK[round == 1 & treatment == "noPunish10", stock],
          #          yend = STOCK[round == 10 & treatment == "noPunish10", stock],
          #          colour = "black", lty = 2, alpha = 0.5) +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
          labs(y = "Wealth") +
          scale_color_manual(values = colors) + 
          theme(plot.margin = margin(0.25,1,0.25,0.25, "cm"))

@@

```{r tableStock, results = "asis", warning = FALSE}

@@

## Inequality

In this subsection, we focus on the degree of inequality created endogenously in our setting. The smallest possible value the Gini coefficient takes is zero (if all four group members own one fourth of the wealth) and the largest possible value it takes is one (if one group member holds the entire wealth). Table 4 shows some summary statistics regarding the Gini coefficient.

The round 10 Gini coefficient ranges between `r main[treatment == "replication" & round == 10, gini %>% min()] %>% round(digits = 3)` and `r main[treatment == "replication" & round == 10, gini %>% max()] %>% round(digits = 3)` in our data with a median of `r main[treatment == "replication" & round == 10, gini %>% median()] %>% round(digits = 3)`.

```{r summary2, results = "asis", warning = FALSE}

replication <- main[treatment == "replication" & round == 10, gini]
GMTV    <- main[treatment == "noPunish10" & round == 10, gini]

stargazer(temp,
          summary.stat = c("mean", "median","sd", "max", "min", "n"),
          type = type, 
          flip = TRUE, 
          header=FALSE)

rs2 <- wilcox.test(replication, 
                   GMTV,
                   exact = FALSE)$p.value %>% 
  round(digits = 4) %>% 
  formatC(format = "f", 
          digits = 4)

The two-sided rank sum test yields a p-Value of `r rs2` for the mean ```gini``` during the last round of the game.

Figure 3 illustrates the dynamics of the Gini coefficient (at the end of each round) over time and shows that inequality increases slightly.

```{r plotInequality, fig.cap = "Average Gini coefficient over time across treatments."}
GINI <- main[,
             lapply(.SD, mean, na.rm = TRUE),
             by = c("round", "treatment"),
             .SDcols = "gini"]

GINIr <- main[rich == TRUE,
              lapply(.SD, mean, na.rm = TRUE),
              by = c("round", "treatment"),
              .SDcols = "gini"]

GINIp <- main[rich == FALSE,
              lapply(.SD, mean, na.rm = TRUE),
              by = c("round", "treatment"),
              .SDcols = "gini"]

upperLimit <- GINI$gini %>% max() %>% round(digits = 1) + 0.15

p1 <- ggplot(data = GINI, 
       aes(x = round, y = gini, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
          labs(y = "Gini Coefficient") +
          scale_color_manual(values = colors) + 
          theme(plot.margin = margin(0.25,1,0.25,0.25, "cm"))

p2 <- ggplot(data = GINIr, 
       aes(x = round, y = gini, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
          labs(y = "Gini (Rich)") +
          scale_color_manual(values = colors)

p3 <- ggplot(data = GINIp, 
       aes(x = round, y = gini, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
          labs(y = "Gini (Poor)") +
          scale_color_manual(values = colors)

rm(list = c("GINI", "GINIp", "GINIr", "p1", "p2", "p3"))
```

### Inequality Differences between Data Sources

Once more, we consider whether we were able to replicate GMTV's results with our data.

The following table shows a simple OLS regression to illustrate differences in the 10th round's Gini coefficient between the replication's data and GMTV's data. Mean Gini coefficients are similar across data sources and there are no statistically significant differences in mean Gini coefficients.

```{r tableGINI, results = "asis", warning = FALSE}

m1 <- lm(formula = gini ~ treatment, data = main_all)
m2 <- lm(formula = gini ~ treatment, data = main_poor)
m3 <- lm(formula = gini ~ treatment, data = main_rich)

          column.labels = c("All", "Below median", "Above median"),
          model.numbers = FALSE,
          dep.var.labels = "Gini",
          header=FALSE,
          covariate.labels = c("Replication"),

# Online Methodology

## Time Spent

Participants spent approximately `r duration[, completion_time %>% unique() %>% mean()] %/% 60 + 1` minutes completing the experiment. Reading the instructions and answering the comprehension questions took the most time, that is, `r duration[page_name == "Intro_Instructions", time_spent %>% unique() %>% mean()] %/% 60 + 1` minutes. The public goods game required `r duration[app_name == "dPGG", time_spent %>% sum(), by = c("session_code", "participant_code")][,V1] %>% mean(na.rm = TRUE) %/% 60` minutes of the participants' time.

```{r plotTime, fig.cap = "Average Time Spent for each Contribution per Round."}
N <- duration[, participant_code %>% unique() %>% length()]
plotDT <- duration[app_name == "dPGG" & page_name == "dPGG_Decision",
                   .(time_spent = time_spent %>% sum()),
                   by = c("session_code", "participant_code", "page_index", "page_name")]

plotDT[, round := seq(from = 1, to = 10), by = c("participant_code")]

upperLimit <- plotDT[, time_spent %>% mean(), by = c("round")] %>% max()

ggplot(data = summarySE(data = plotDT,
                        measurevar = "time_spent",
                        groupvars=c("round"),
                        na.rm = FALSE,
                        conf.interval = 0.95,
                        .drop = TRUE),
       mapping = aes(x = round, y = time_spent)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE, color = cSecondary, lty=2) +
  geom_errorbar(aes(ymin=time_spent-ci, ymax=time_spent+ci), width=.25, alpha = 0.5, color = cSecondary) +
  geom_point(color = cSecondary) +
  scale_x_continuous(name="",  breaks = 1:10) +
  scale_y_continuous(limits = c(0, upperLimit + 10), expand = c(0, 0)) +
  labs(y = "Time Passed in Seconds", caption = "Bars indicate 95% confidence intervals.") +
  theme(plot.margin = margin(0.25,1,0.25,0.25, "cm"))
```

Figure 4 illustrates the time participants needed to make each of their contributions during the replication. One can see that the initial as well as most of the other contributions take about 20 seconds of time. Interestingly, the second contribution takes (on average) about 50% more time than the first one - presumably because this is the first time participants' learn their respective group members' actions. The third contribution is a little faster and all subsequent contributions stabilize at `r plotDT[round > 3, time_spent %>% mean()] %>% ceiling()` seconds.

Given that no participant dropped out after answering the comprehension questions correctly and given that participants need less than 20 seconds to make their contributions, more than 10 rounds are feasible.

# Exporatory: Generalizability

The following figure gives a first overview over the distribution of contributions across the two sorts of contributions. The density-plots in the top panel show the fraction of the initial endowment contributed in the first round of the experiment. The two plots in the bottom panel show contribution behavior in the VCA, that is, donations to climate change mitigation. The left panels depict the general population's behavior while the right panelshows the students' behavior. All contributions are relative to the maximum amount a participant could have contributed.

Means are indicated by vertical dashed lines and are positive in both settings across samples.^[While the medians are similar in the bottom panel, that is, the experiment's contributions, they are smaller in the taop panel and zero in the student sample.] Overall contributions in the VCA setting are smaller than in the experiment's first round.

```{r vizGeneralizability}
R1covs <- covs[, .(participant.code,
                   donation = donation * 20 ,
                   donationShare = (donation * 20) / payoff * 100,
                   payoff)]
temp <- R1[R1covs, on = .(participant.code = participant.code)][treatment == "replication"]
temp[, contributionShare := ownContribution/20 * 100]

# viz wrapper
plotShares <- function(stud = FALSE,
                       var  = "donationShare",
                       col  = cPrimary,
                       xlab = "",
                       ylab = ""){

  dt <- temp[student == stud, .(x = get(var))]

  ggplot(data = dt,
         mapping = aes(x = x)) +
    geom_density(alpha = 1,
                 fill = col) +
    scale_x_continuous(limits = c(0, 100),
                       expand = c(0, NA)) +
    scale_y_continuous(limits = c(0, NA),
                       expand = c(0, NA)) +
    geom_vline(xintercept = dt[,mean(x, na.rm = TRUE)],
               lty = 2) +
    labs(x = xlab,
         y = ylab) +
  layout
}

p1 <- plotShares(stud = FALSE, col = cPrimary,   var = "donationShare", ylab = "Density in VCA")
p2 <- plotShares(stud = FALSE, col = cPrimary,   var = "contributionShare", 
                 xlab = "Share of general pop. contributions", ylab = "Density in Experiment (round 1)")
p3 <- plotShares(stud = TRUE,  col = cSecondary, var = "donationShare")
p4 <- plotShares(stud = TRUE,  col = cSecondary, var = "contributionShare", xlab = "Share of students' contributions")

(p1 / p2) | (p3 / p4)

rm(list = c("R1covs", "p1", "p2", "p3", "p4"))
```

lorem ipsum.

```{r}
ggplot(data = temp[, 
                   .(vca = donation / payoff * 100,
                     dpgg = ownContribution / 20 * 100)],
       mapping = aes(y = vca, x = dpgg)) +
  annotate("segment", x = 0, xend = 100,
                   y = 0,
                   yend = 100,
                   colour = "black", lty = 2, alpha = 0.5) +
  geom_point(alpha = 0.33, col = cSecondary, size = 2) +
  scale_x_continuous(limits = c(0, 105),
                     expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, 105),
                     expand = c(0, NA)) +
  labs(y = "VCA: Percentage of endowment donated",
       x = "Experiment: Percentage of endowment contributed in round 1") +
  layout
```
A visual inspection of the scatter plot (where the saturation of the points is proportional to the frequency of a pair of contribution choices) does not hint at an association between the two sorts of contributions.

## To do

- explain similarities and differences to @GKLS2020 setup and how they came to be
- compare samples across tasks
- who generalizes (better)?
- who is more heterogeneous and what does that mean?
- How do the results compare to @GKLS2020

# Conclusion

>>>>

---
title: Growth and inequality in public good provision | an extended replication
author:
  - name: Hauke Roggenkamp
    email: hauke.roggenkamp@unisg.ch
    affiliations:
        - id: HSU
          name: Helmut Schmidt University
          department: Department of Economics
          address:  Holstenhofweg 85
          city: Hamburg
          state: Germany
          postal-code: 22043
        - id: HSG
          name: University of St. Gallen
          department: Institute for Behavioral Science and Technology
          address: Torstrasse 25
          city: St. Gallen
          state: Switzerland
          postal-code: 9000
    # attributes:
    #     corresponding: true
    # note: This is the first author footnote.
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - Replication study
  - Non-convenience sample
  - Open science
  - Dynamic public good game
  - Online experiment
  - Generalizability
date: last-modified
bibliography:  ../biblio.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal of Behavioral and Experimental Economics
      formatting: preprint
      model: 3p
      cite-style: authoryear
---

@@

```{r treatmentAsFactor}
main[, treatment := as.factor(treatment)]
R1[, treatment := as.factor(treatment)]
covs[, treatment := as.factor(treatment)]
```

@@

> "There are two possible articles you can write: (a) the article you planned to write when you designed your study or (b) the article that makes the most sense now that you have seen the results. They are rarely the same, and the correct answer is (b)." [@bemwriting, p. 171]

@@

@@

First, we ask whether the samples differ with respect to their initial contributions to the public good. Is our replication sample more pro-social than the original sample? Figure 1 reveals, that it is not. The distributions of both samples look fairly similar. Both samples contributed 10 tokens, that is, 50% of their endowments on average (median and mean).^[The two-sided rank sum test (comparing differences between samples) yields a p-Value of `r rs1` for the mean contribution in first round of the game.] Moreover, both samples' initial contributions resemble initial contributions participants usually make in the standard game with partner matching.^[See Figure 3B in @fehrgaechter2000 (p.989), for instance.] In the dynamic game presented here, we are particularly interested in the subsequent periods because differences add up exponentially. Do the two groups remain similar?

```{r firstRoundViz}
ggplot(data = R1[treatment != "noPunish15"],
       mapping = aes(x = ownContribution, fill = treatment, lty = treatment)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(0, 20),
                       expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, 0.1),
                       expand = c(0, NA)) +
  scale_fill_manual(values = c(cPrimary, cSecondary),
                    labels = c("Original Sample", "Replication Sample")) +
  guides(lty = "none") +
  geom_vline(xintercept = R1[treatment == "replication", 
                             mean(ownContribution)],
             col = "#FFFFFF",
             lty = 2) +
  geom_vline(xintercept = R1[treatment == "noPunish10", 
                             mean(ownContribution)],
             col = "#FFFFFF",
             lty = 1) +
    labs(title = "Individual contributions to the dynamic public good in the first period", 
       y = "", x = "Contributrions (Tokens)",
       caption = "White lines indicate means (dashed line = replication sample).") +
  layout +
  theme(legend.position = "bottom")
```

```{r firstRoundSummary}
#| results: asis
#| warning: false
#| fig.cap: "test"
#| eval: false

@@

In particular, do the two samples' contributions follow the same pattern over time? The answer is _no_. Figure 2 illustrates that the samples make similar contributions at the beginning and the end of the game but behave differently in between. More precisely, the left panel--depicting the average contributions in absolute terms--shows that the original sample contributed substantially more than the replication sample _in all but the first and last rounds_. For this reason, the original sample's behavior differs from the replication sample's behavior in two aspects: it contributes more and exhibits a considerable drop in the last period (whereas the replication sample's contributions flatten). Note that increasing contributions over time imply increasing endowments over time. Hence, increasing absolute contributions do not necessarily imply that participants contribute increasing shares of their endowments. The right panel in Figure 1 shows the share of overall endowments contributed over time where both samples exhibit a similar pattern: they decline and do not stabilize. The original sample's behavior differs from the replication sample's behavior in one aspect: it declined slower (to the same level in the last period).

@@

SUM[,
    contribution := contribution/4]

upperLimit <- SUM$contribution %>% max() %>% round() + 5

p1 <- ggplot(data = SUM, 
             aes(x = round, y = contribution, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Average Amount of Tokens contributed") +
  scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) + 
  guides(fill = "none") +
  theme(plot.margin = margin(0.25,1,0.25,0.25, "cm"))

@@

```{r plotShareOfContributions}
#| fig.cap: "The average amount of tokens contributed over time in treatments."

@@

p2 <- ggplot(data = SHARE, 
             aes(x = round, y = share, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Share of Current Endowment contributed") +
  scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) +
  guides(fill = "none")

@@

# rm(list = c("p1", "p2"))
```

Moreover, both samples' contributions resemble contributions participants usually make in the standard game with partner matching.^[The right panel is thus, comparable to the visualizations _and results_ in the standard game. See, for instance, Figure 1B in @fehrgaechter2000 (p.986).] In the dynamic game presented here, different paths lead to different levels of wealth -- even if they share the same start- and end-points. We are thus, more interested in the implications contributions have for wealth generation and growth. So, how do the samples compare with respect to wealth/income/endowments/stock?

@@

```{r summary1}
#| results: asis
#| warning: false
#| fig.cap: "test"

@@

# stargazer(temp,
#           summary.stat = c("mean", "median","sd", "max", "min", "n"),
#           type = type, 
#           flip = TRUE, 
#           header=FALSE)

@@

How do the different contribution-paths translate to wealth?^[To measure growth, we define a variable called _stock_ which sums the endowments of all participants in a given group at the end of the round (that is, after the contributions have been made, multiplied and redistributed).] Given that the original sample contributed more in most of the periods, one would expect the respective groups to be considerably more wealthy. Figure 3 indicates that they tend to be insignificantly^[The two-sided rank sum test (comparing differences between samples) yields a p-Value of `r rs1` for the mean stock in last round of the game.] richer.

```{r stockDistributionViz}
originalWealth    <- main[treatment == "noPunish10" & round == 10, mean(stock)]
replicationWealth <- main[treatment == "replication" & round == 10, mean(stock)]

ggplot(data = main,
       mapping = aes(x = stock, fill = treatment, lty = treatment)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(0, NA),
                       expand = c(0, NA)) +
  scale_y_continuous(limits = c(0, NA),
                       expand = c(0, NA)) +
  scale_fill_manual(values = c(cPrimary, cSecondary),
                    labels = c("Original Sample", "Replication Sample")) +
  guides(lty = "none") +
  geom_vline(xintercept = replicationWealth,
             col = "#000000", alpha = 0.5,
             lty = 2) +
  geom_vline(xintercept = originalWealth,
             col = "#000000", alpha = 0.5,
             lty = 1) +
    labs(title = "Groups' income at the end of the game", 
       y = "", x = "Stock",
       caption = "Grey lines indicate means (dashed line = replication sample).") +
  layout +
  theme(legend.position = "bottom")
```

As the grey lines indicate, an average group in the original sample accumulated about `r round(originalWealth)` tokens. In contrast, an average group in the replication sample accumulated about `r round(replicationWealth)` tokens. However, none of the groups were efficient: the maximal wealth that can be reached in round 10 under full cooperation is approximately `r (80*1.5^10) %>% trunc()` tokens or `r (80*1.5^10 /20) %>% trunc()` Euro per group.

```{r growthHeterogeneityViz}
maxPath <- data.table(x = 1:10)
maxPath[, y := 80*1.5^x]
maxPath[, groupID := 42]
ggplot(data = main,
       mapping = aes(x = round, y = stock, color = treatment, by = groupID)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(alpha = 0.5) +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0, 4700),
                       expand = c(0, NA)) +
  scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) +
  geom_point(mapping = aes(x = 10, y = 4613), col = "#b5b5b5") +
  annotate("text", x = 9.7, y = 4630,
  label = "max", col = "#b5b5b5") +
  geom_line(data = maxPath, mapping = aes(x = x, y =y), col = "#b5b5b5", lty = 2)
```

While there is clearly growth, groups do not realize the potential efficiency as the replication groups reach on average a level of `r main[treatment == "replication" & round == 10, mean(stock)] %>% trunc()` tokens out of `r (80*1.5^10) %>% trunc()` maximally possible or `r main[treatment == "replication" & round == 10, mean(stock)/(80*1.5^10)*100] %>% round(digits=1)`%. As in the original data, there is large heterogeneity with the richest group reaching `r main[treatment == "replication" & round == 10, max(stock)]` tokens whereas the poorest group ends up with `r main[treatment == "replication" & round == 10, min(stock)]` tokens.

@@

Figure 2 shows the dynamics of wealth over time.^[Unlike GMTV I present no median split but only the main results.]

```{r plotStock}
#| fig.cap: "Average wealth over time across treatments."
STOCK <- main[,
              lapply(.SD, mean, na.rm = TRUE),
              by = c("round", "treatment"),
              .SDcols = "stock"]

@@

p1 <- ggplot(data = STOCK, 
       aes(x = round, y = stock, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          # annotate("segment", x = 1, xend = 10, 
          #          y = STOCK[round == 1 & treatment == "replication", stock],
          #          yend = STOCK[round == 10 & treatment == "replication", stock],
          #          colour = "black", lty = 2, alpha = 0.5) +
          # annotate("segment", x = 1, xend = 10, 
          #          y = STOCK[round == 1 & treatment == "noPunish10", stock],
          #          yend = STOCK[round == 10 & treatment == "noPunish10", stock],
          #          colour = "black", lty = 2, alpha = 0.5) +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, 600), expand = c(0, 0)) +
          labs(y = "Wealth") +
          scale_color_manual(values = c(cPrimary, cSecondary),
                     labels = c("Original Sample", "Replication Sample")) +
          guides(fill = "none") +
          theme(plot.margin = margin(0.25,1,0.25,0.25, "cm"))

p1
```

```{r plotStockLikeGMTV}
#| eval: false

@@

```{r tableStock}
#| results: asis
#| warning: false

@@

### 2022-10-08 analysis/reports/00_reproduction.Rmd
added "quarto" and "paper" (already deprecated) + no clue what happened otherwise. Its been a long time

In an attempt to incorporate uncertainty to @GACHTER2017's dynamic public goods game (DPGG), I plan to run a series of remote online experiments using oTree [@oTree]. The first experiment will replicate Gächter et al.'s NOPUNISH 10-period version as close as possible (given the remote circumstances). The current demo version of the experiment can be found [here](https://cliccs.herokuapp.com/). Click [here](https://github.com/Howquez/coopUncertainty) to visit the corresponding Github repository.

This document explores the data provided by @GACHTER2017 and finds that some of the variables are a little off. In addition, it fails to replicate the orginal article's figures. The following chapter will describe the available data. Subsequently, [data excerpts](#Tables) will demonstrate that the original _Gini coefficient_ varies within groups (in selected periods) and the _share of current endowment contributed_ is calculated using the last round's endowment. [Afterwards](#Downloads), the data is processed and made available for downloads.The [Visualizations](#Visualizations) chapter will then present the original figures as well as figures reproduced using the original and re-calculated variables to demonstrate differences.

>>>>

In an attempt to incorporate uncertainty to @GMTV2017's dynamic public goods game (DPGG), I plan to run a series of remote online experiments using oTree [@oTree]. The first experiment will replicate Gächter et al.'s NOPUNISH 10-period version as close as possible (given the remote circumstances). The current demo version of the experiment can be found [here](https://cliccs.herokuapp.com/). Click [here](https://github.com/Howquez/coopUncertainty) to visit the corresponding Github repository.

This document explores the data provided by @GMTV2017 and finds that some of the variables are a little off. In addition, it fails to replicate the orginal article's figures. The following chapter will describe the available data. Subsequently, [data excerpts](#Tables) will demonstrate that the original _Gini coefficient_ varies within groups (in selected periods) and the _share of current endowment contributed_ is calculated using the last round's endowment. [Afterwards](#Downloads), the data is processed and made available for downloads.The [Visualizations](#Visualizations) chapter will then present the original figures as well as figures reproduced using the original and re-calculated variables to demonstrate differences.
